Spatial Analysis

### 2021-08-19 Scripts/spatial_analisis.Rmd
Fixed grid for matching extrapolation

# us_state_file <- my_path("G", extra_path = "Spatial/USFedAndStateW/FederalAndStateWaters.gdb")

state_sf <-  st_read("~/OneDrive - UW-Madison/Spatial_Data/USFedAndStateW/FederalAndStateWaters.gdb") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  mutate(state = str_to_sentence(jurisdicti)) %>% 
  st_transform(crs = 4326) %>%   # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) #0.1 for paper

@@

# Buffer state waters
port_bf = st_buffer(port_sf, 5) %>% # port
  st_transform(4326) %>% # to match shape
  st_set_crs(4326)

@@

@@

# Get the overlapping segments 
# NOTE: Takes some gooooood time (~20 min)
Sys.time()
grid_eez_sf_sw <- st_intersection(grid_sw_sf,eez_sf)
Sys.time()
# write_sf(grid_eez_sf, paste0(my_path("D", "Spatial"),"grid_eez_sf.shp"))

@@

# write_csv(grid_eez_df, paste0(my_path("R", "Partial"),"grid_eez_df.csv"))

@@

# write_sf(grid_eez_sf_fp, paste0(my_path("D", "Spatial/grid_eez_fp_sf"),"grid_eez_fp.shp"))

@@

write_csv(grid_eez_fp_df, paste0(my_path("R", "Partial"),"grid_eez_fp_df.csv"))

@@

grid_eez_sw_df %>% 
  mutate(approach = "state_waters") %>% 
  bind_rows(grid_eez_fp_df) %>%
  mutate(approach = ifelse(is.na(approach),"fishing_port",approach)) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      color = approach,
      fill = approach
    )
  ) +
  facet_wrap(~state) +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2") +
  theme_dark()

@@

# Get state order from North to south
state_order <- my_path("D","Spatial/grid_eez_fp_sf", name = "grid_eez_fp_df.csv",read = T) %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(order= c("b","a",letters[3:11])) %>% # weirdly maine goes below
  arrange(order)

# Load grid of state waters
grid_eez_sw_sf <- st_read(my_path("D","Spatial/grid_eez_sw_sf", name = "grid_eez_sf.shp")) %>% 
  select(index,state,geometry) %>% 
  mutate(method = "state_waters",
         state = str_to_sentence(state)) %>% 
  left_join(state_order)

@@

# Load grid of fishing ports
grid_eez_fp_sf <- st_read(my_path("D","Spatial/grid_eez_fp_sf", name = "grid_eez_fp.shp")) %>% 
  select(index,state = jursdct,geometry) %>% 
  mutate(method = "fishing_port",
         state = str_to_sentence(state))%>% 
  left_join(state_order)

@@

ggsave(filename = "buffer_figure.jpg",
       plot = combined_plot,
       path = my_path("R","Figures"),
       width = 11,
       height = 11
)

@@

- We removed cases where `wtcpue < 0`

@@

# Load grid df
# For fishing ports
grid_eez_fp_df <- my_path("D","Spatial/grid_eez_fp_sf","grid_eez_fp_df.csv", read = T)

# For state waters
# grid_eez_sw_df <- my_path("D","Spatial/grid_eez_fp_sf","grid_eez_fp_df.csv", read = T)

@@

# Run them all in parallel
lapply(spp,
       run_tis, 
       input_data = ocean_data,
       grid = grid_eez_fp_df,
       years = years,
       reg = regions,
       method = "fp"
)

@@

@@

# Load interpolated data
# "Centropristis striata"
# 
# historic_tif <- my_path("R","Partial/Interpolation","tif_Centropristis_striata.csv", read = T) %>% 
#   mutate(state = str_to_sentence(state))

# Multiple species

@@

Species <- c(
  # "Scomber scombrus",
  # "Doryteuthis (Amerigo) pealeii", #longfin squid WE DONT HAVE IT
  # "Peprilus triacanthus", # butterfish
  # "Illex illecebrosus", #shortfin squid
  "Paralichthys dentatus", #summer flounder
  "Stenotomus chrysops", #Scup"
  "Centropristis striata"#, # black sea bass
  # "Pomatomus saltatrix", # Bluefish
  # "Lopholatilus chamaeleonticeps", #golden tilefish
  # "Caulolatilus microps", #blueline tilefish
  # "Clupea harengus" # Atlantic herring (Jointly in State and Federal Waters)
  )

@@

taxon_fp <- my_path("R",extra_path = "Partial/Interpolation/fp", name = taxon_read)

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  mutate(spatial = "sw") %>% 
  bind_rows(
    bind_rows(lapply(taxon_fp, fread))
  ) %>% 
  mutate(state = str_to_sentence(state),
         spatial = ifelse(is.na(spatial),"fp",spatial))

unique(historic_spp$spp)
unique(historic_spp$spatial)

@@

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_spp_spatial.jpg",
  p,
  base_height = 14,
  base_width = 14
)

@@

state_fit <- historic_spp %>% 
  group_by(state,year,region,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spatial")) %>%
  mutate(percentage = state_value/total_value*100, 
  # left_join(periods,
            # by = "year") %>% 
  label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  group_by(state,label,region,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  filter(str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  spread(spatial,mean_per) %>% 
  mutate(difference = (fp-sw)/((fp+sw)/2)*100,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# Spatial map
agg_sppatial_map <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Distribution\nProportion", alpha = 0.8) +
  facet_grid(spatial~ label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "right",
               leg_tl_s = 14,
               ax_tx_s = 10,
               ax_tl_s = 14) +
theme(legend.title.align=0.5)

# Difference map
agg_diff_map <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Percentage\nDifference",option = "C", alpha = 0.8) +
  facet_grid(spatial~ label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "right",
               leg_tl_s = 14,
               ax_tx_s = 10,
               ax_tl_s = 14)+
  theme(
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.title.align=0.5
        )

# Cowplot option
p <- ggdraw() +
  # Revenue circular
  draw_plot(agg_sppatial_map, x = 0, y = 0.35, width = 1, height = 0.65) +
  # Catch circular
  draw_plot(agg_diff_map, x = 0.093, y = 0.05, width = 0.814, height = 0.4) +
  draw_plot_label(label = c("Latitude", "Longitude"), 
                  size = 10,
                  angle = c(90,0),
                  x = c(0,0.35), 
                  y = c(0.45,0.11)
  )

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_agg.jpg",
  p,
  base_height = 8,
  base_width = 8
)

>>>>

state_sf <-  st_read("~/OneDrive - UW-Madison/FederalAndStateWaters/FederalAndStateWaters.gdb") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  mutate(state = str_to_sentence(jurisdicti)) %>% 
  st_transform(crs = 4326) %>%   # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) #0.1 for paper

@@

# Buffer state waters
port_bf <- st_buffer(port_sf, 5) %>% # port
  st_transform(4326) %>% # to match shape
  st_set_crs(4326)

@@

##### 2.1 Create grid for interpolation

```{r inter_grid, eval = T, echo = T}

# Crop grid to EEz
inter_grid_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_intersection(eez_sf)

inter_grid_df <- as.data.frame(inter_grid_sf) %>% 
  select(index) %>% 
  left_join(ne_grid)

write_csv(inter_grid_df, paste0(my_path("D", "Spatial"),"inter_grid_df.csv"))

ggplot(inter_grid_sf) +
  geom_sf()

@@

# Get the overlapping segments 
# NOTE: Takes some gooooood time (~20 min)
Sys.time()
grid_eez_sw_sf <- st_intersection(grid_sw_sf,eez_sf)
Sys.time()

write_sf(grid_eez_sw_sf, paste0(my_path("D", "Spatial/grid_eez_sw_sf"),"grid_eez_sw_sf.shp"))

@@

write_csv(grid_eez_sw_df, paste0(my_path("D", "Spatial/grid_eez_sw_sf","grid_eez_sw_df.csv")))

@@

write_sf(grid_eez_sf_fp, paste0(my_path("D", "Spatial/grid_eez_fp_sf"),"grid_eez_fp.shp"))

@@

write_csv(grid_eez_fp_df, paste0(my_path("D", "Spatial/grid_eez_fp_sf"),"grid_eez_fp_df.csv"))

@@

grid_eez_sw_df %>% 
  mutate(approach = "state_waters") %>% 
  bind_rows(grid_eez_fp_df) %>%
  mutate(approach = ifelse(is.na(approach),"fishing_port",approach)) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = approach
    ),
    alpha = 0.5
  ) +
  facet_wrap(~state) +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2") +
  theme_dark()

@@

# Get state order from North to south
state_order <- my_path("D","Spatial/grid_eez_sw_sf", name = "grid_eez_sw_df.csv",read = T) %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(order= c("b","a",letters[3:11])) %>% # weirdly maine goes below
  arrange(order)

# Load grid of state waters
grid_eez_sw_sf <- grid_eez_sw_sf %>% 
  # st_read(my_path("D","Spatial/grid_eez_sw_sf", name = "grid_eez_sw_sf.shp")) %>%
  select(index,state,geometry) %>% 
  mutate(method = "state_waters",
         state = str_to_sentence(state)) %>% 
  left_join(state_order)

@@

# Load grid of fishing ports
grid_eez_fp_sf <- grid_eez_sf_fp %>% 
  # st_read(my_path("D","Spatial/grid_eez_fp_sf", name = "grid_eez_fp.shp")) %>% 
  select(index,state,geometry) %>% 
  mutate(method = "fishing_port",
         state = str_to_sentence(state))%>% 
  left_join(state_order)

@@

ggsave(filename = "buffer_figureb.jpg",
       plot = combined_plot,
       path = my_path("R","Figures"),
       width = 11,
       height = 11
)

@@

@@

# Load grid df
# For fishing ports
inter_grid <- my_path("D","Spatial","grid_eez_fp_df.csv", read = T)

@@

# Run them all in parallel
lapply(species,
       run_tis, 
       input_data = ocean_data,
       grid = grid_eez_fp_df,
       years = years,
       reg = regions,
       method = "fp"
)

@@

### Previouslly created grids

grids <- my_path("D", "Spatial/grid_eez_fp_sf", name = "grid_eez_fp_df.csv", read = T) %>% 
  mutate(
    spatial = "fp"
  ) %>% 
  bind_rows(
    my_path("D", "Spatial/grid_eez_sw_sf", name = "grid_eez_sw_df.csv", read = T)
  ) %>% 
  mutate(
    spatial = ifelse(is.na(spatial),"sw",spatial)
  ) %>% 
  spread(spatial,index)

@@

@@

Species <- c(
  "Paralichthys dentatus", #summer flounder
  "Stenotomus chrysops", #Scup"
  "Centropristis striata"#, # black sea bass
  )

@@

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  mutate(state = str_to_sentence(state)) %>% 
  # dplyr::select(-lon,-lat) %>% 
  full_join(grids,
            by = c("index","state")
            )

historic_spp %>% 
  filter(is.na(spatial)) %>% 
  filter(year == 1973,
         region == "Northeast US Fall") %>% 
  ggplot() +
    geom_tile(
      aes(
        x = lon.x,
        y = lat.x,
        fill = index
      )
    ) +
  facet_grid(spp~state) + 
  geom_sf(data = land_sf,aes()) +
  geom_sf(data = eez_sf, aes(), fill =NA)

unique(historic_spp$spp)
unique(historic_spp$spatial)
unique(historic_spp$region)

@@

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_spp_spatialN.jpg",
  p,
  base_height = 14,
  base_width = 14
)

@@

state_change <- historic_spp %>% 
  group_by(state,year,region,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"reference",
                         ifelse(year > 2001,"today",NA)
                         )
         ) %>% 
  # left_join(periods,
            # by = "year") %>% 
  group_by(state,label,region,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(#str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  # select(-region) %>% 
  # Estimate percentage change
  spread(label,mean_per) %>% 
  mutate(change = (today-reference)/(today)*100,
         change = ifelse(change > 100,100,
                         ifelse(change < -100,-100,change)
         )
  ) %>%
  select(-reference,-today)

  # Estimate spatial differences in percentage change
season_diff <- state_change %>% 
  spread(spatial,change) %>% 
  mutate(difference_season = (fp-sw)/((fp+sw)/2)*100,
         difference_season = ifelse(difference_season > 100,100,
                                  ifelse(difference_season < -100,-100,difference_season)
         )
  ) %>% 
  select(state, category = region, diff = difference_season)

# Estimate regional differences in percentage change
spatial_diff <- state_change %>% 
  spread(region,change) %>%  
  mutate(difference_spat = (`Northeast US Fall`-`Northeast US Spring`)/((`Northeast US Fall`+`Northeast US Spring`)/2)*100,
         difference_spat = ifelse(difference_spat > 100,100,
                                 ifelse(difference_spat < -100,-100,difference_spat)
         )
  ) %>% 
  select(state,category = spatial,diff = difference_spat)

differences <- season_diff %>% 
  bind_rows(spatial_diff) %>% 
  mutate(
    label = ifelse(category == "fp","Fishing ports",
                          ifelse(category == "sw","State Waters",category)
         )
  )

# The main plot
prop_chng_map <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  # filter(spatial != "Difference" & cat != "difference_reg") %>% 
  ggplot() +
  geom_sf(aes(fill = change)) +
  viridis::scale_fill_viridis("Change in proportion today\n relative to 1973-1984\n(%)", alpha = 0.8,
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c(seq(-100,75,25),">100")
                              ) +
  facet_grid(spatial~ region) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5,
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        )

ggsave(filename = "proportion_chng_agg.jpg",
         plot = prop_chng_map,
         path = my_path("R","Figures"),
         width = 10,
         height = 10
  )

# Season differencce

spat_diff <- land_sf %>% 
  left_join(differences,
            by = "state") %>% 
  filter(!category %in% c("fp","sw")) %>% 
  ggplot() +
  geom_sf(aes(fill = diff)) +
  viridis::scale_fill_viridis("Differences", 
                              alpha = 0.8,
                              option = "C",
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c(seq(-100,75,25),">100")
                              ) +
  facet_wrap(~ label,
             ncol = 2) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5,
        strip.background = element_blank(),
        strip.text.x = element_blank())

ggsave(filename = "spat_diff.jpg",
         plot = spat_diff,
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )

# seasonal
seas_diff <- land_sf %>% 
  left_join(differences,
            by = "state") %>% 
  filter(category %in% c("fp","sw")) %>% 
  ggplot() +
  geom_sf(aes(fill = diff)) +
  viridis::scale_fill_viridis("Differences", 
                              alpha = 0.8,
                              option = "C",
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c(seq(-100,75,25),">100")
                              ) +
  facet_wrap(~ label,
             ncol = 1,
             strip.position="right") +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(strip.background = element_blank(),
        # strip.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank()
        )

ggsave(filename = "seas_diff.jpg",
         plot = seas_diff,
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )
